---
title: "DATA607 | Project 1 "
author: "Ait Elmouden Abdellah"
date: "02/22/2020"
output:
  html_document:
    df_print: paged
---


### Objective of the Assignment:

The Objective of the this project is to processes and extract information from a text file that includes chess tournament results and create a R script that generates an .CSV file, after extracting the needed information.

### Data source 

The provided chess tournament results text file was uploaded to Github, and it will be loaded from there.

### Load the raw Data from Github :

Lets load the data from the text file as a character vector:

```{r}
tournamentdata <- readLines('https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/project1/tournamentinfo.txt')
```

Lets Check the head of the data:

```{r}
head(tournamentdata, 16)
```
##Clean up the data##

The stringr package provide a cohesive set of functions designed to make working with strings as easy as possible.

```{r}
library(stringr)
```

### Remove the dashes using sub function and Regex ###

```{r}
nodash <- sub("^-+$", "", tournamentdata) # This also can be done using str_replace_all function
tail(nodash)
```

### Clean the empty rows ###


```{r}
noblanklines <- nodash[sapply(nodash, nchar) > 0]
head(noblanklines)
```

### Generate regular sequences to create separate vector tables ###

**Players Names Table**

```{r}
PlayerName <- noblanklines[seq(1, 130, 2)]
head(PlayerName)  
```

**Players ID's Table**

** Same function will be used to extract the player ID)

```{r}
PlayerID <- noblanklines[seq(2, 130, 2)]
head(PlayerID)
```

**Remove header rows from PlayerName and PlayerID**

```{r}
PlayerNameNoHeader <- PlayerName[-c(1:1)]
head(PlayerNameNoHeader)
```

```{r}
PlayerIDNoHeader <- PlayerID[-c(1:1)]
head(PlayerIDNoHeader)
```

**make all the data about one player combined in one row**

```{r}
Combinedata <- mapply(paste, sep = "", PlayerNameNoHeader, PlayerIDNoHeader)
head(Combinedata)
```

###  Extract the relevant information from the txt data using regular expressions 

**Extract USCFID**

```{r}
USCFID <- str_extract(string = Combinedata, pattern = "[0-9]{8}")
USCFID
```

**Extract Player ID**

```{r}
PlayerID <- str_extract(string = Combinedata, pattern = "[\\s{3}]\\d{1,2}[\\s\\|]")
PlayerID <- str_trim(PlayerID)
PlayerID
```

**Extract player Names**

```{r}
Name <- str_extract(string = Combinedata, pattern = "(\\w+\\s){2,5}")
Name <- str_trim(Name)
Name
```

**Extract Player States**

```{r}
State <- str_extract(string = PlayerIDNoHeader, pattern = "[[:alpha:]]{2}")
State
```

**Extract the Total Number of Points**

```{r}
NbrPoints <- str_extract(string = Combinedata, pattern = "[0-9]\\.[0-9]")
NbrPoints
```

**Extract the Player’s Pre-Rating**

```{r}
PreRating <- str_extract(string = Combinedata, pattern = "\\s\\d{3,4}[^\\d]")
PreRating <- as.integer(str_extract(PreRating, "\\d+")) #Remove all the “P’s” at the end of the ratings
PreRating
```

### Calculate Average Pre Chess Rating of Opponents


*In this step we'll extract the Opponents data by Player ID*


```{r}
opponentData <- unlist(str_extract_all(PlayerNameNoHeader, "([\\|][A-Z]([[:space:]]+)\\d*[\\|])([A-Z]([[:space:]]+)\\d*[\\|])*", simplify=TRUE))
opponents <- matrix(ncol=7)
opponentData
```

**Get the individual Opponent Indexes into a matrix of 7 columns**

```{r}
opponents <- unlist(str_extract_all(opponentData[,], "\\d+", simplify=TRUE))
head(opponents)
```

**In this step we'll Loop through each row of Opponent Index. Match each Opponent Index with its corresponding Pre-Rating. Get the average Opponent rating for each row**

```{r}
RatingAverages = NULL #initialize teh rating averages 

for(row in 1:nrow(opponents)){
  numberOfgames = 0
  sum = 0
  
  for(col in 1:ncol(opponents)){
    
    if(opponents[row, col] != ""){ # Check to make sure we are not looking at a null opponent index value
      Oppindex <- opponents[row, col] # Get the Opponent Index
      Oppindex <- strtoi(Oppindex, base=0L) # Convert to integer
      sum = sum + strtoi(PreRating[Oppindex]) # Update sum of corresponding pre-ratings
      numberOfgames = numberOfgames + 1 # Update number of opponents
    }
  }
  
  avgerage = sum/numberOfgames
  RatingAverages = round(rbind(RatingAverages, data.frame(avgerage)),digits = 0)
}
head(RatingAverages)
```

## Create a data frame with the relevant information

```{r}
ChessResults <- data.frame(PlayerID, Name, State, NbrPoints, PreRating, RatingAverages)
head(ChessResults, 5)
```

**Change the column names**

```{r}
colnames(ChessResults) <- c("PlayerID","Player's Name", "State", "Total Number of Points", "Pre-Rating", "Average Rating of Opponents")
head(ChessResults, 5)
```

## Generates a .CSV file

```{r}
write.csv(ChessResults, file = "ChessResults.csv")
```

[Uplease the file to Github](https://github.com/aaitelmouden/DATA607/blob/master/ChessResults.csv)

### References ###

"Handling and Processing Strings in R", [Gaston Sanchez](https://www.gastonsanchez.com/)
http://www.datasciencemadesimple.com/seq-function-in-r/
https://stackoverflow.com/questions/37901142/paste-multiple-rows-together-in-r
Wickham, H. (2017). R: [Package stringr. Retrieved February 12, 2019, from]( https://cran.r-project.org/web/packages/stringr/stringr.pdf)
