---
title: "DATA607 | Week 5"
author: "Ait Elmouden Abdellah"
date: "02/29/2020"
output:
  html_document:
    df_print: paged
---


### Objective of the Assignment:

The goal of this assignment is to use R to tidy (reshape) data, using "base R" functionality and Hadley Wickham's tidyr and dplyr packages. 
The folowing data that was used in this assignment describes arrival delays for two airlines across five destinations. 

|         |         | Los.Angeles | Phoenix | San.Diego | San.Franciso | Seattle |
|---------|---------|-------------|---------|-----------|--------------|---------|
| ALASKA  | On time | 497         | 221     | 212       | 503          | 1841    |
|         | Delayed | 62          | 12      | 20        | 102          | 305     |
|         |         |             |         |           |              |         |
| AM WEST | On Time | 694         | 4840    | 383       | 320          | 201     |
|         | Delayed | 117         | 415     | 65        | 129          | 61      |


*Data Source : Numbersense, Kaiser Fung, McGraw Hill, 2013*


### Load the raw Data from Github :

Lets load the data from the text file as a character vector:

```{r}
tournamentdata <- readLines('https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/project1/tournamentinfo.txt')
```

Lets Check the head of the data:

```{r}
head(tournamentdata, 16)
```
##Clean up the data##

The stringr package provide a cohesive set of functions designed to make working with strings as easy as possible.

```{r}
library(stringr)
```

### Remove the dashes using sub function and Regex ###

```{r}
nodash <- sub("^-+$", "", tournamentdata) # This also can be done using str_replace_all function
tail(nodash)
```

### Clean the empty rows ###


```{r}
noblanklines <- nodash[sapply(nodash, nchar) > 0]
head(noblanklines)
```

### Generate regular sequences to create separate vector tables ###

**Players Names Table**

```{r}
PlayerName <- noblanklines[seq(1, 130, 2)]
head(PlayerName)  
```

**Players ID's Table**

** Same function will be used to extract the player ID)

```{r}
PlayerID <- noblanklines[seq(2, 130, 2)]
head(PlayerID)
```

**Remove header rows from PlayerName and PlayerID**

```{r}
PlayerNameNoHeader <- PlayerName[-c(1:1)]
head(PlayerNameNoHeader)
```

```{r}
PlayerIDNoHeader <- PlayerID[-c(1:1)]
head(PlayerIDNoHeader)
```

**make all the data about one player combined in one row**

```{r}
Combinedata <- mapply(paste, sep = "", PlayerNameNoHeader, PlayerIDNoHeader)
head(Combinedata)
```

###  Extract the relevant information from the txt data using regular expressions 

**Extract USCFID**

```{r}
USCFID <- str_extract(string = Combinedata, pattern = "[0-9]{8}")
USCFID
```

**Extract Player ID**

```{r}
PlayerID <- str_extract(string = Combinedata, pattern = "[\\s{3}]\\d{1,2}[\\s\\|]")
PlayerID <- str_trim(PlayerID)
head(PlayerID)
```

**Extract player Names**

```{r}
Name <- str_extract(string = Combinedata, pattern = "(\\w+\\s){2,5}")
Name <- str_trim(Name)
head(Name)
```

**Extract Player States**

```{r}
State <- str_extract(string = PlayerIDNoHeader, pattern = "[[:alpha:]]{2}")
head(State)
```

**Extract the Total Number of Points**

```{r}
NbrPoints <- str_extract(string = Combinedata, pattern = "[0-9]\\.[0-9]")
head(NbrPoints)
```

**Extract the Player’s Pre-Rating**

```{r}
PreRating <- str_extract(string = Combinedata, pattern = "\\s\\d{3,4}[^\\d]")
PreRating <- as.integer(str_extract(PreRating, "\\d+")) #Remove all the “P’s” at the end of the ratings
head(PreRating)
```

### Calculate Average Pre Chess Rating of Opponents


*In this step we'll extract the Opponents data by Player ID*


```{r}
opponentData <- unlist(str_extract_all(PlayerNameNoHeader, "([\\|][A-Z]([[:space:]]+)\\d*[\\|])([A-Z]([[:space:]]+)\\d*[\\|])*", simplify=TRUE))
opponents <- matrix(ncol=7)
head(opponentData)
```

**Get the individual Opponent Indexes into a matrix of 7 columns**

```{r}
opponents <- unlist(str_extract_all(opponentData[,], "\\d+", simplify=TRUE))
head(opponents)
```

**In this step we'll Loop through each row of Opponent Index. Match each Opponent Index with its corresponding Pre-Rating. Get the average Opponent rating for each row**

```{r}
RatingAverages = NULL #initialize teh rating averages 

for(row in 1:nrow(opponents)){
  numberOfgames = 0
  sum = 0
  
  for(col in 1:ncol(opponents)){
    
    if(opponents[row, col] != ""){ # Check to make sure we are not looking at a null opponent index value
      Oppindex <- opponents[row, col] # Get the Opponent Index
      Oppindex <- strtoi(Oppindex, base=0L) # Convert to integer
      sum = sum + strtoi(PreRating[Oppindex]) # Update sum of corresponding pre-ratings
      numberOfgames = numberOfgames + 1 # Update number of opponents
    }
  }
  
  avgerage = sum/numberOfgames
  RatingAverages = round(rbind(RatingAverages, data.frame(avgerage)),digits = 0)
}
head(RatingAverages)
```

## Create a data frame with the relevant information

```{r}
ChessResults <- data.frame(PlayerID, Name, State, NbrPoints, PreRating, RatingAverages)
head(ChessResults, 5)
```

**Change the column names**

```{r}
colnames(ChessResults) <- c("PlayerID","Player's Name", "State", "Total Number of Points", "Pre-Rating", "Average Rating of Opponents")
head(ChessResults, 5)
```

## Generates a .CSV file

```{r}
write.csv(ChessResults, file = "ChessResults.csv")
```

[Final CVS file in Github](https://github.com/aaitelmouden/DATA607S2020/blob/master/project1/ChessResults.csv)

### References ###

1. "Handling and Processing Strings in R", [Gaston Sanchez](https://www.gastonsanchez.com/)
2. http://www.datasciencemadesimple.com/seq-function-in-r/
3. https://stackoverflow.com/questions/37901142/paste-multiple-rows-together-in-r
4. Wickham, H. (2017). R: [Package stringr. Retrieved February 12, 2019, from]( https://cran.r-project.org/web/packages/stringr/stringr.pdf)
