---
title: "DATA607 | Week 7"
author: "Ait Elmouden Abdellah"
date: "03/13/2020"
output:
  html_document:
    df_print: paged
---


## Objective of the Assignment:
The goal of this assignment is to manipulate Less-structured data from the web. We'll focus on the standard formats for web data:  HTML, XML, and JSON.  Three books were selected and information were stored in three files HTML, XML and JASON, from where data will be loaded.

## Load the required Libraries 

```{r}
library(XML) #for xml processing
library(rvest)  # to scrape (or harvest) data from html web pages
library(jsonlite) # for json files processing
library("kableExtra") # Construct Complex Table with 'kable' and Pipe Syntax

```


## HTML file Manupulation

```{r}
htmlfile <- "https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/Week7/book.html"

htmlTable <- read_html(htmlfile)
htmlBooks <- htmlTable %>%
  html_nodes("table") %>%  # Select nodes from THE HTML document
  .[[1]] %>%
  html_table(fill = NA) # converts data to an R data frame automatically

colnames(htmlBooks) <- c("Category","Title", "Authros", "Publisher","Published Date", "ISBN")

htmlBooks  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")

```

## XML file Manupulation

```{r}
xmlfile <- "https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/Week7/book.xml"
 
xmlData <- read_xml(xmlfile)
Books <- xml_children(xmlData)

xmlBooks <- c()
for (i in 1:length(Books)){
  xmlBooks <- rbind(xmlBooks,xml_text(xml_children(Books[i])))
}

xmlBooks <- data.frame(xmlBooks)

colnames(xmlBooks) <- xml_name(xml_children(Books[1]))

xmlBooks  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")


```

## JSON Manupulation

```{r}
jsonfile <- "https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/Week7/book.json"

jsonBooks <- fromJSON(jsonfile) #converts any json object to an R data frame.

jsonBooks  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")

```

## Conclusion

While all 3 files html, xml and json have different structures , We were able to get the same resulting dataframes As we can see.