---
title: "DATA607 | Week1 Assignment"
author: "Ait Elmouden Abdellah"
date: "01/30/2020"
output:
  html_document:
    df_print: paged
---
### Objective of the Assignment:
The goal of this assignment is to practice Loading Data into a Data Frame, study the dataset and the associated description of the data. The data was available from fivethirtyeight.com datasets. the selected data is behind an article about [Picking A College Major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/). 
All data is from American Community Survey 2010-2012 Public Use Microdata Series. , it was uploaded to [github](https://github.com/aaitelmouden/DATA607S2020/blob/master/Week1/recent-grads.csv).

### Data Preparation:

*First we will load the data from Github*

```{r}

CollegeMajors <- read.csv("https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/Week1/recent-grads.csv")

```

*Obtain the first several rows : Let’s view the dataset before we start transforming it:*
```{r}
head(CollegeMajors)
# Retrieve the dimension of the Data.
dim(CollegeMajors)
```

```{r}
# Column Names in Our Data Frame: names column names are meaningful, so no replacment is needed. instead headers for recent-grads.csv are shown below:
```


| Header               | Description                                                                |
|----------------------|----------------------------------------------------------------------------|
| Major_code           | Rank by median earnings                                                    |
| Rank                 | Major code, FO1DP in ACS PUMS                                              |
| Major                | Major description                                                          |
| Major_category       | Category of major from Carnevale et al                                     |
| Total                | Total number of people with major                                          |
| Sample_size          | Sample size (unweighted) of full-time, year-round ONLY (used for earnings) |
| Men                  | Male graduates                                                             |
| Women                | Female graduates                                                           |
| ShareWomen           | Women as share of total                                                    |
| Employed             | Number employed (ESR == 1 or 2)                                            |
| Full_time            | Employed 35 hours or more                                                  |
| Part_time            | Employed less than 35 hours                                                |
| Full_time_year_round | Employed at least 50 weeks (WKW == 1) and at least 35 hours (WKHP >= 35)   |
| Unemployed           | Number unemployed (ESR == 3)                                               |
| Unemployment_rate    | Unemployed / (Unemployed + Employed)                                       |
| Median               | Median earnings of full-time, year-round workers                           |
| P25th                | 25th percentile of earnigns                                                |
| P75th                | 75th percentile of earnings                                                |
| College_jobs         | Number with job requiring a college degree                                 |
| Non_college_jobs     | Number with job not requiring a college degree                             |
| Low_wage_jobs        | Number in low-wage service jobs                                            |

### Create a data frame with a subset of the columns in the dataset:

```{r}
MajorEmploy <- subset(CollegeMajors, select = c(Major, Employed, Unemployed, Median))

head(MajorEmploy)
```


### Data Explorations & Graphics

```{r}
# We'll use the level function to have access to the available Majors.
levels(CollegeMajors$Major)
```
### Salaries

```{r}
library(tidyverse)
theme_set(theme_light())
CollegeMajors %>% # pipe operation
  ggplot(aes(Median)) + geom_histogram()
```

So as we can seen most major's make a little over 30K

### Most common Majors

```{r}
CollegeMajors %>%
  count(Major_category, wt = Total, sort = TRUE) %>%
  ggplot(aes(Major_category, n)) + geom_col() + coord_flip() +
labs(x = "",
     y = "Total # of Graduates")
```

## Conclusion 

R allows practicing a wide variety of statistical and graphical techniques. Using R I was able to explore dataset about college Majors, and after reading a little bit about tidyverse library, I was able to produced a handful of figures that I found pretty interesting. 