---
title: "DATA607 | Week 5"
author: "Ait Elmouden Abdellah"
date: "02/29/2020"
output:
  html_document:
    df_print: paged
---


### Objective of the Assignment:

The goal of this assignment is to use R to tidy (reshape) data, using "base R" functionality and Hadley Wickham's tidyr and dplyr packages. 
The folowing data that was used in this assignment describes arrival delays for two airlines across five destinations. 

|         |         | Los.Angeles | Phoenix | San.Diego | San.Franciso | Seattle |
|---------|---------|-------------|---------|-----------|--------------|---------|
| ALASKA  | On time | 497         | 221     | 212       | 503          | 1841    |
|         | Delayed | 62          | 12      | 20        | 102          | 305     |
|         |         |             |         |           |              |         |
| AM WEST | On Time | 694         | 4840    | 383       | 320          | 201     |
|         | Delayed | 117         | 415     | 65        | 129          | 61      |


*Data Source : Numbersense, Kaiser Fung, McGraw Hill, 2013*

### Load the required Libraries 

```{r}
library("tidyr")
library("dplyr")
```

We will load the data from csv file hosted in github:

```{r}
airlines <- read.csv("Link")
```

Lets Check the head of the data:

```{r}
head(tournamentdata, 16)
```
